---
title: "STA 141a Final Project"
author: "Brett Loy"
date: "2023-08-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Abstract(what you did on the project(3-4 pages))
Introduction
- questions of interest:
Will Stephen Curry's record of 402 3 pointers made in a single season ever be broken?
- information of data source
- existing research and known results
- impact of results along with real world

Exploratory Analysis(describe the data)
-basic description statistics(missing values(what to do with them can replace with mean or median))
-data visualization(show features/patterns in data)

Methodology
-Ex) predictive model(how you apply statistical model into your data)
  1)fit model
  2)report fitted model
 
Result
-prediction performance and evaluate it

Discussion, Conclusion
-main funding, next steps


what are the chances of that nba players that will pass lebron james in points?
factors: time spent in league(games and years played), playstyle changes(3 pt line, overall shift in focus on offense, potential 4 point line)

always things that you can't predict: injuries, adjustments, bench players coming in and out of rotation,
trades?

disclaimer: data is only from 1999-2020 . Three pointers are more common now
filter data so it only contains data from 1999-2023/combine datasets together
```{r}
library(readr)

#read data in
nba_data <- #1999-2020
read_csv("/Users/brettloy/Desktop/STA 141A/Datasets/players_stats_by_season_full_details.csv")

nba_extra_shooting <- #1997-2023
  read_csv("/Users/brettloy/Desktop/STA 141A/Datasets/Player Shooting.csv")
```


```{r}
library(dplyr)
library(stringr)
#wrangle the data to end with 1 data frame to work with
years_filter_out <- c("1997", "1998", "1999", "2021", "2022", "2023")

nba_extra_shooting<- nba_extra_shooting %>% 
  #choose which columns from the data set needed to keep
  select("season", "player", "fg_percent_from_x3p_range", "percent_assisted_x3p_fg",
         "corner_3_point_percent", "percent_corner_3s_of_3pa")%>%
  #filter out seasons that aren't needed
  filter(!(season %in% years_filter_out))%>%
  #rename columns
  rename("percent_of_3PM_assisted"="percent_assisted_x3p_fg", "three_PM_percentage"=
"fg_percent_from_x3p_range", "percent_of_3PA_incorner"="percent_corner_3s_of_3pa", 
"percent_of_3PM_incorner" = "corner_3_point_percent", "Season" = "season", "Player"="player")

#regular expression, subset, to keep the last 4 digits of season
nba_data<- nba_data %>%
  #choose which columns from the data set needed to keep
  select("League", "Season", "Player", "GP", "MIN", "3PM", "3PA", "PTS" ) %>%
  #filter out 
  filter(League=="NBA")%>%
  rename("three_point_makes" = "3PM", "three_point_attempts" = "3PA")
nba_data$Season <- str_extract(nba_data$Season, "\\d{4}$")
#change "Season" column to character in order to join
nba_extra_shooting$Season<- as.character(nba_extra_shooting$Season)
#merge data frames sorting by "Season" and "Player"
join_df <- left_join(nba_data, nba_extra_shooting, by = c("Season", "Player"))

join_df
```

```{r}
library(ggplot2) 
library(ggthemes)
library(ggrepel)
library(ggsci)

curry_df <- join_df %>% filter(Season ==2016) %>% 
  filter(Player == "Stephen Curry") %>%
  mutate(threes_per_game = three_point_makes/GP)
curry_df #Record: 402 3PM, 886 3PA
```


```{r}
#3 point makes per game of players
plot_mper_game<- join_df %>% filter(three_point_makes>=250) %>%
  mutate(threes_mper_game = three_point_makes/GP) %>%
  ggplot(aes(threes_mper_game, Player, label = Season))+
  geom_label()+
  ggtitle("3 Pointers Made per Game by Players who've come close to the record")+
  xlab("3 Pointers Made per Game")+
  theme_clean()
plot_mper_game
#3 point attempts per game of players
plot_aper_game<- join_df %>% filter(three_point_makes>=250) %>%
  mutate(threes_aper_game = three_point_attempts/GP) %>%
  ggplot(aes(threes_aper_game, Player, label = Season))+
  geom_label()+
  ggtitle("3 Pointers Attempted per Game by Players who've come close to the record")+
  xlab("3 Pointers Attempted per Game")+
  theme_clean()
plot_aper_game
```


```{r}
#players who have come close

plot_close<- join_df %>% filter(three_point_makes>=250) %>%
  arrange(desc(three_point_makes))%>%
  ggplot(aes(three_point_makes, Player, label = Season, color = Player))+
  geom_label()+
  ggtitle("Players in the Past Who've Come Close to the Record")+
  xlab("3 Pointers Made")+
  theme_clean()
plot_close

avg_3pt_per <- join_df %>% group_by(Season)%>% filter(three_point_makes>=275)%>%
  ggplot(aes(Season, three_PM_percentage, label = Player, color = Player))+
  ggtitle("3 Point Percentages of players who've made more than 275 Threes")+
  geom_label()+
  theme_clean()
avg_3pt_per

corner_3pt_per <- join_df %>% group_by(Season)%>% filter(three_point_makes>=270)%>%
  ggplot(aes(Season, percent_of_3PM_incorner, label = Player, color = Player))+
  ggtitle("3 Pt Percentages from Corners of players who've made more than 270 Threes")+
  geom_label()+
  theme_clean()
corner_3pt_per
```


```{r}
#data shows the increase in 3s made and attempted throughout the years
filtered_season1 <- subset(join_df, Season %in% c(2000, 2005, 2010, 2015, 2020))

three_attempt_2000_2020<- filtered_season1 %>% 
  ggplot(aes(three_point_attempts))+
  geom_bar(width = 6, color = "darkblue", fill = "blue")+
  ggtitle("3 Point Shots Attempted in 2000, 2005, 2010, 2015, and 2020")+
  xlab("3 Pointers Attempted")+
  ylab("Number of Players")+
  facet_wrap(~Season)+
  theme_clean()
three_attempt_2000_2020

three_made_2000_2020<- filtered_season1 %>% 
  ggplot(aes(three_point_makes))+
  geom_bar(width = 6, color = "darkred", fill = "red")+
  ggtitle("3 Point Shots Made in 2000, 2005, 2010, 2015, and 2020")+
  xlab("3 Pointers Made")+
  ylab("Number of Players")+
  facet_wrap(~Season)+
  theme_clean()
three_made_2000_2020


```


```{r}
avg_data_makes <- join_df %>%
  group_by(Season) %>%
  summarize(Avg_3PM = mean(three_point_makes)) %>%
  ggplot(aes(Season, Avg_3PM))+
  geom_point()+
  ggtitle("Average 3 Point Makes from 2000-2020")+
  ylab("Average number of 3 pt Makes")+
  theme_clean()
avg_data_makes

avg_data_attempts <- join_df %>%
  group_by(Season) %>%
  summarize(Avg_3PA = mean(three_point_attempts)) %>%
  ggplot(aes(Season, Avg_3PA))+
  geom_point()+
  ggtitle("Average 3 Point Attempts from 2000-2020")+
  ylab("Average number of 3 pt Attempts")+
  theme_clean()
avg_data_attempts

```

```{r}
join_df%>% filter(Season>= 2015)%>% 
  mutate(threes_aper_game = three_point_makes/GP)%>% 
  
predict_model <- lm(three_point_makes~three_PM_percentage + threes_aper_game, data = join_df )

```

  
  
  
  
  
  
  
  